---
title: "Creation of bacteremia Analysis File"
author: "Barry DeCicco"
date: "2023-05-31"
output: html_document:
    self-contained: true
    code-link: false
    code-tools: true
    code-fold: hide

---

```{r}
require(Hmisc)
require(qreport)
options(prType='html')
require(rio)
d <- import('https://zenodo.org/record/7554815/files/bacteremia-DataDictionary.csv?download=1')
d <- subset(d, Variable != 'BloodCulture')
v <- tolower(d$Variable)
labels <- d$Label
units  <- d$Units
names(labels) <- v
names(units)  <- v
units['id'] <- units['sex'] <- ''

labels <- c(labels, bacteremia='Bacteremia present by blood culture')
units  <- c(units,  bacteremia='')
labels['glu'  ] <- 'Glucose'
labels['baso' ] <- 'Basophils'
labels['ck'   ] <- 'Creatine kinase'
labels['basor'] <- 'Basophil ratio'

d <- import('https://zenodo.org/record/7554815/files/Bacteremia_public_S2.csv?download=1')
names(d) <- tolower(names(d))
d <- upData(d,
            sex        = factor(sex, 1:2, c('male', 'female')),
            bacteremia = 1L * (bloodculture == 'yes'),
            labels = labels,
            units  = units,
            drop   = 'bloodculture')
options(prType='html')

Hmisc::contents(d)
```

Click on the tabs to see the different kinds of variables.  Hover over spike histograms to see frequencies and details about binning.

```{r results='asis'}
bacteremia <- d
sparkline::sparkline(0)   # load javascript dependencies for interactive spike histograms

maketabs(print(describe(bacteremia), 'both'), cwidth='column-screen-inset-shaded')
```

```{r}
save(bacteremia, file='bacteremia.sav', compress=TRUE)
cat(html(contents(bacteremia)), file='Cbacteremia.html')
```

Here is the older html output format for `describe`.

```{r}
describe(bacteremia)
```

## Data Overview

```{r results='asis'}
dataOverview(bacteremia, id = ~ id)
```

## Missing Data Report

```{r results='asis'}
missChk(bacteremia, prednmiss=TRUE, omitpred='id')
```

## Variable Clustering Report

The R `Hmisc` package `transace` function, which uses the ACE (alternating conditional expectation) algorithm, is used to transform all the continuous variables.  Transformations use nonparametric smoothers and are allowed to be non-monotonic.  Transformation solutions maximize the $R^2$ which with each variable can be predicted from the other variables, optimally transformed.  The transformed variables are used in redundancy analysis and sparse principal components analysis.  Bacteremia and subject `id` are not used in these unsupervised learning procedures.

To be more efficient, use multiple (5) imputations with predictive mean matching so that `vClus` will stack all the filled-in datasets before running the redundancy and PCA which are run on the single tall dataset, which contains no `NA`s.  The correlation matrix and `varclus` results are already efficient because they use pairwise deletion of `NA`s.

Because transformed variables are passed to the redundancy analysis, variables are not expanded into splines in that analysis (see `nk=0` below).


```{r results='asis'}
n <- setdiff(names(bacteremia), 'id')
n[n == 'baso'] <- 'I(baso)'  # force baso to be linear in MI b/c of ties
f <- as.formula(paste('~', paste(n, collapse='+')))
midone <- TRUE  # took aregImpute about 15 minutes
if(! midone) {
  set.seed(1)   # so that mult. imputations reproduce
  a <- aregImpute(f, data=bacteremia, n.impute=5)
  saveRDS(a, 'a.rds')
  } else a <- readRDS('a.rds')
v <- vClus(bacteremia, fracmiss=0.8, corrmatrix=TRUE,
           trans=TRUE, redundancy=TRUE, spc=TRUE,
           exclude = ~ id + bacteremia,
                     imputed=a,
                 redunargs=list(nk=0),
                 spcargs=list(k=20, sw=TRUE, nvmax=5), # spc took 5 minutes
                     transacefile='transace.rds',
                 spcfile='spc.rds')   # spcfile uses previous run if nothing changed
cat('R^2 in predicting each variable with transace:\n\n')
htmlVerbatim(round(attr(v$transace, 'rsq'), 3))
saveRDS(v, '/tmp/v.rds')
```

```{r}
#| fig-height: 12
#| fig-width: 7
ggplot_transace(v$transace, nrow=12)
```

```{r}
#| fig-height: 4
#| fig-width: 6
p <- v$princmp
# Print and plot sparse PC results
print(p)
plot(p)
```

```{r}
#| fig-height: 7
#| fig-width: 8
#| column: screen-right
plot(v$p, 'loadings', nrow=1)
```
